Experimentally informed site-specific substitution models deepen phylogenetic estimates of the divergence of viral lineages

### 2018-02-19 manuscript/manuscript.tex
updates to references

Accounting for site-to-site rate variation in GY94 models affects the rate at which the asymptotic sequence identity is approached, but not the actual value of this asymptote. 
For instance, \ref{fig:decay} shows that the GY94+$\Gamma\omega$ model takes longer to reach the asymptote than GY94, but the asymptote for both models is identical. 
This fact holds true even if we use experimental measurements of HA's site-specific amino-acid preferences~\citep{doud2016accurate} to calculate a different $\omega_r$ value for each site using the method of \citet[\ref{eq:w_r};][]{spielman2015relationship}.
Specifically, this GY94+$\omega_r$ model predicts that different sites will approach the asymptote at different rates, but the asymptote is always the same (\ref{fig:decay}).
The invariance of the asymptotic sequence identity under different schemes for modeling $\omega$ is a fundamental feature of the mathematics of homogeneous substitution models.
The asymptotic sequence divergence is a function of only the stationary state of the substitution matrix, and the stationary state of a stochastic matrix is invariant with respect to multiplication by any non-zero number\jdbcomment{But this isn't true, because $\omega$ only multiplies some elements of transition matrix. I think it's probably something like if you multiple pairs of elements on equivalent positions across diagonal?}.
Therefore, different schemes for modeling $\omega$ can alter the time it takes to reach stationary state, but no matter how ``well" a model accounts for site-to-site variation in $\omega$ it will always have the same stationary state as a simple GY94 model.

\jdbcomment{In contrast, ExpCM models have different stationary states at each site...}
\skhcomment{Google books: reversible matrix exchangeability stationary state}

>>>>

Accounting for site-to-site rate variation in GY94 models affects the rate at which the asymptotic sequence identity is approached, but not the actual value of this asymptote. 
For instance, \ref{fig:decay} shows that the GY94+$\Gamma\omega$ model takes longer to reach the asymptote than GY94, but the asymptote for both models is identical. 
This fact holds true even if we use experimental measurements of HA's site-specific amino-acid preferences~\citep{doud2016accurate} to calculate a different $\omega_r$ value for each site using the method of \citet[\ref{eq:w_r};][]{spielman2015relationship}.
Specifically, this GY94+$\omega_r$ model predicts that different sites will approach the asymptote at different rates, but the asymptote is always the same (\ref{fig:decay}).
The invariance of the asymptotic sequence identity under different schemes for modeling $\omega$ is a fundamental feature of the mathematics of homogeneous substitution models.
The asymptotic sequence divergence is a function of only the stationary state of the substitution matrix, and the stationary state of a stochastic matrix is invariant with respect to multiplication by any non-zero number~\citep{nielsen2006statistical}.
Therefore, different schemes for modeling $\omega$ can alter the time it takes to reach stationary state, but no matter how ``well" a model accounts for site-to-site variation in $\omega$ it will always have the same stationary state as a simple GY94 model.

\jdbcomment{In contrast, ExpCM models have different stationary states at each site...}
